Applying the metallicity-dependent binary fraction to double white dwarf formation: \\ Implications for LISA

### 2022-08-19 src/tex/ms.tex
minor text edits

\begin{table}[]
    \centering
    \begin{tabular}{|c|c|c|c|c|c|c|}
       \hline
       parameter variation & model & a & b & c & d & e \\
       \hline
       \hline
       fiducial & F50 & -217.8 & -183.9 & -74.5 & -13.7 & -0.94452821 \\
       \hline
       fiducial & FZ & -268.2 & -243.0 & -100.5 & -18.7 & -1.3 \\
       \hline
       $\alpha25$ & F50 & -5520.2 & -6078.3 & -2528.8 & -467.1 & -32.3 \\
       \hline
       $\alpha25$ & FZ & -2816.0 & -3066.4 & -1272.7 & -234.7 & -16.2 \\
       \hline
       $\alpha5$ & F50 & -178.7 & -142.8 & -58.7 & -11.0 & -0.8 \\
       \hline
       $\alpha5$ & FZ & -265.9 & -239.8 & -99.1 & -18.5 & -1.3 \\
       \hline
       q3 & F50 & -114.5 & -73.5 & -30.9 & -6.1 & -0.5 \\
       \hline
       q3 & FZ & -336.9 & -323.5 & -136.2 & -25.8 & -1.8 \\
       \hline
    \end{tabular}
    \caption{Polynomial fitting coefficients for the confusion foreground fit of Equation~\ref{eq:fit} for each binary fraction model.}
    \label{tbl:fits}
\end{table}

@@

To prevent the two stars' envelopes from being stripped before helium ignition, CO + CO DWDs typically form from progenitors in wider orbits, and the two components may have little to no interaction during their evolution from ZAMS to DWD. CO DWDs thus have a distribution in progenitor separation that extends to larger values than for other DWD types. Most need $> 0.3$ Gyr to form. These systems have component WD masses between $0.35$--$1.0\,M_\odot$ and make up the majority of the DWD population.

ONe WDs are rare and typically form from massive progenitor stars which evolve through the asymptotic giant branch (AGB) phase, thus resulting in a higher mass WD. All ONe WDs in our \cosmic\ populations, e.g., have progenitor ZAMS masses above $4\,M_\odot$, and the resulting ONe WDs have a relatively flat distribution of masses from $1.05\,M_\odot$ up to the Chandrasekkhar limit of $1.4\,M_\odot$. Because an ONe WD can have a companion of any other WD type in our study, there is a spread in their distributions for separation, secondary mass, final orbital period, and formation time. In general, however, these systems result from wider separations to allow for the evolution of the ONe component without merging, e.g. all initial separations in our \cosmic \ populations have separations $\gtrsim 1.5\,R_\odot$. ONe + X DWDs can form on short timescales, as low as $30\,\rm{Myr}$ for the majority of high-metallicity systems.

\subsection{Metallicity-dependent trends in the formation efficiency of DWDs}\label{sec:formeff}
The number of DWDs formed per unit solar mass of ZAMS star formation, or DWD formation efficiency $\eta_{\rm{form}}(Z)$, varies with metallicity. Consequently, a metallicity-dependent binary fraction further impacts the efficiency of DWD formation within the Galaxy. Figure \ref{fig:form_eff} shows the DWD formation efficiency as a function of metallicity for each DWD type\snotes{,\sout{ and for each of our}} binary fraction models\snotes{, and binary evolution parameter variations}. In general, the formation efficiency decreases with increasing metallicity. This effect is exaggerated for model FZ which assumes a binary fraction which also decreases with increasing metallicity.

@@

For He + He DWDs, the sudden drop in formation efficiency near $\log(Z/Z_{\odot})=-1.0$, is generally caused by the timescale for which the initially more massive star in the DWD progenitor overflows it's Roche lobe. At lower metallicities, donors tend to fill their Roche lobes while they are still on the main sequence and the mass transfer remains stable \snotes{and serves to widen the binary}. This is because, in our \cosmic\ models lower metallicity stars evolve faster than high-metallicity stars and have larger radii near the end of the MS. At higher metallicities, mass transfer is initiated when the donor has left the main sequence, and the binary enters a common envelope evolution. The ZAMS orbital period leads to different specific evolutionary channels. For short-period systems with periods below $10~\rm{days}$, this leads to a stellar merger due to insufficient orbital energy to eject the envelope. Stellar mergers continue to dominate the evolutionary pathways of systems with intermediate orbital periods ($1 < \log_{10}(P_{\rm{orb}}/\rm{day}) < 2.5$). However, an additional growing number of merging systems arises with one WD component and one stellar companion, and systems which don't interact at all and thus do not form a He + He DWD. The distinction between the various scenarios in this intermediate orbital period range depends on the combination of their ZAMS masses and orbital periods. Finally, at wider initial periods $\log_{10}(\rm{P}) > 2.5$, the decrease in formation efficiency is dominated by systems which never interact and thus do not form a DWD before the present day.

@@

For short-period CO + He DWD progenitor binaries with orbital periods below $\sim30~\rm{days}$, the only channel for mergers before DWD formation is during a CE evolution. This channel is similar to the stellar merger channel for He + He DWDs, but due to higher-mass progenitors the CE evolution results in the binary component with a higher mass becoming a CO WD. At higher metallicities, the CO WD merges with its companion. Higher-metallicity progenitors have a larger fraction of their mass in the convective envelope when compared to lower-metallicity stars of the same mass \citep{Amard2019, Amard2020}. Thus a CE phase with a higher-metallicity progenitor requires more orbital energy to eject the common envelope, causing a larger amount of orbital shrinking which results in a merger later in its binary evolution.

For CO + He DWD progenitor systems with intermediate orbital periods ($1.5\leq \log_{10}(P_{\rm{orb}}/\rm{day})< 2.5$) the mechanisms which impact formation efficiency are complex. The dominant way to lose CO + He DWDs is stellar mergers that occur during the second CE phase. For lower-metallicity systems, the second CE is successful, while the opposite is true for higher-metallicity systems. This is again because the envelope mass of the CE evolution donor at higher metallicity is larger than it would be at lower metallicity and thus requires more orbital energy to eject in the CE phase. This leads to shorter post-CE orbital periods after the first CE phase and hence mergers during the second phase.

@@

Long-period binaries with $\log_{10}(P_{\rm{orb}}/\rm{day})>2.5$ also display complex scenarios that hinder CO + He DWD formation. In near-equal contributions, our \cosmic\ simulations produce either stellar mergers or stable non-DWD binaries at the end of the Hubble time. At these orbital periods, stellar mergers always occur with a CE phase between a stellar companion and a CO or He WD. Similar to binaries with shorter orbital periods, the mergers occur because of increased CE donor envelope masses at higher metallicities. A subdominant channel of stable He + He DWDs can also occur when a high-metallicity primary overflows it's Roche lobe while still on the GB and thus forms a He WD. \snotes{In general, a CO + He binaries fails to form because either a CO + CO or He + He is formed isntead, or a merger occurs before DWD formation. We see higher efficiency for FZ than F50 because Co + He binaries prefer to form in log$_{10}$(P$_{\rm{orb}}$/day)$\sim2-4$, which is where the orbital period distribution of FZ peaks.}

@@

\snotes{For our simulations with parameter variations, we see an interplay of the physics of binary evolution with the skews introduced by the metallicity-dependent orbital period distribution/binary fraction. We will summarize how the formation efficiency of each DWD type is impacted by the parameter variations. Of note in this analysis is that a) the log-normal metallicity-dependent ZAMS orbital period distribution increasingly skews towards orbital periods below $10^4$ days for decreasing metallicity. However, b) the log-uniform P$_{\rm{orb}}$ distribution of F50 produces a higher number of the shortest orbital period systems (below $\sim 100$ days) than for FZ as the tails of the FZ distribution fall quickly below this threshold.}

>>>>

\begin{table}[]
    \centering
    \begin{tabular}{|c|c|c|c|c|c|}
       \hline
       model & a & b & c & d & e \\
       \hline
       \hline
       fiducial, F50 & -217.8 & -183.9 & -74.5 & -13.7 & -1.0 \\
       \hline
       fiducial, FZ & -268.2 & -243.0 & -100.5 & -18.7 & -1.3 \\
       \hline
       $\alpha25$, F50 & -5520.2 & -6078.3 & -2528.8 & -467.1 & -32.3 \\
       \hline
       $\alpha25$, FZ & -2816.0 & -3066.4 & -1272.7 & -234.7 & -16.2 \\
       \hline
       $\alpha5$, F50 & -178.7 & -142.8 & -58.7 & -11.0 & -0.8 \\
       \hline
       $\alpha5$, FZ & -265.9 & -239.8 & -99.1 & -18.5 & -1.3 \\
       \hline
       q3, F50 & -114.5 & -73.5 & -30.9 & -6.1 & -0.5 \\
       \hline
       q3, FZ & -336.9 & -323.5 & -136.2 & -25.8 & -1.8 \\
       \hline
    \end{tabular}
    \caption{Polynomial fitting coefficients for the confusion foreground fit of Equation~\ref{eq:fit} for each binary fraction model and parameter variation.}
    \label{tbl:fits}
\end{table}

@@

To prevent the two stars' envelopes from being stripped before helium ignition, CO + CO DWDs typically form from progenitors in wider orbits, and the two components may have little to no interaction during their evolution from ZAMS to DWD. CO DWDs thus have a distribution in progenitor separation that extends to larger values than for other DWD types. Most \kb{CO + CO} DWDs need $> 0.3$ Gyr to form\kb{\sout{. These systems},} have component WD masses between $0.35$--$1.0\,M_\odot$, and make up the majority of the DWD population.

ONe WDs are rare and typically form from massive progenitor stars which evolve through the asymptotic giant branch (AGB) phase, thus resulting in a higher mass WD. All ONe WDs in our \cosmic\ populations, e.g., have progenitor ZAMS masses above $4\,M_\odot$, and the resulting ONe WDs have a relatively flat distribution of masses from $1.05\,M_\odot$ up to the Chandrasekkhar limit of $1.4\,M_\odot$. Because an ONe WD can have a companion of any other WD type in our study, there is a spread in their distributions for separation, secondary mass, final orbital period, and formation time. In general, however, these systems result from wider separations to allow for the evolution of the ONe component without merging\kb{\sout{, e.g.}. For example,} all initial separations in our \cosmic \ populations have separations $\gtrsim 1.5\,R_\odot$. ONe + X DWDs can form on short timescales, as low as $30\,\rm{Myr}$, for the majority of high-metallicity systems.

\subsection{Metallicity-dependent trends in the formation efficiency of DWDs}\label{sec:formeff}
The number of DWDs formed per unit solar mass of ZAMS star formation, or DWD formation efficiency $\eta_{\rm{form}}(Z)$, varies with metallicity. Consequently, a metallicity-dependent binary fraction further impacts the efficiency of DWD formation within the Galaxy. Figure \ref{fig:form_eff} shows the DWD formation efficiency as a function of metallicity for each DWD type\snotes{,\sout{ and for each of our}} binary fraction models\snotes{, and binary evolution parameter variations}. In general, the formation efficiency decreases with increasing metallicity. This effect is exaggerated for model FZ which assumes a \kb{close} binary fraction which also decreases with increasing metallicity. \kb{In the following sub-sections, we describe the formation efficiency trends for our fiducial model in detail and then describe the changes arising from each binary evolution parameter variation.}

@@

For He + He DWDs, the sudden drop in formation efficiency near $\log(Z/Z_{\odot})=-1.0$, is generally caused by the timescale for which the initially more massive star in the DWD progenitor overflows it's Roche lobe. At lower metallicities, donors tend to fill their Roche lobes while they are still on the main sequence and the mass transfer remains stable \snotes{and serves to widen the binary}. This is because, in our \cosmic\ models lower metallicity stars evolve faster than high-metallicity stars and have larger radii near the end of the MS. At higher metallicities, mass transfer is initiated when the donor has left the main sequence, and the binary enters a common envelope evolution. The ZAMS orbital period leads to different specific evolutionary channels. For short-period systems with periods below $10~\rm{days}$, \kb{\sout{this} and common envelope} leads to a stellar merger due to insufficient orbital energy to eject the envelope. Stellar mergers continue to dominate the evolutionary pathways of systems with intermediate orbital periods ($1 < \log_{10}(P_{\rm{orb}}/\rm{day}) < 2.5$). However, an additional growing number of merging systems arises with one WD component and one stellar companion, and systems which don't interact at all and thus do not form a He + He DWD. The distinction between the various scenarios in this intermediate orbital period range depends on the combination of their ZAMS masses and orbital periods. Finally, at wider initial periods $\log_{10}(\rm{P}) > 2.5$, the decrease in formation efficiency is dominated by systems which never interact and thus do not form a DWD before the present day.

@@

For short-period CO + He DWD progenitor binaries with orbital periods below $\sim30~\rm{days}$, the only channel for \kb{stellar} mergers before DWD formation is during a CE evolution. This channel is similar to the stellar merger channel for He + He DWDs, but due to higher-mass progenitors the CE evolution results in the binary component with a higher mass becoming a CO WD. At higher metallicities, the CO WD merges with its companion. Higher-metallicity progenitors have a larger fraction of their mass in the convective envelope when compared to lower-metallicity stars of the same mass \citep{Amard2019, Amard2020}. Thus a CE phase with a higher-metallicity progenitor requires more orbital energy to eject the common envelope, causing a larger amount of orbital shrinking which results in a merger later in its binary evolution.

For CO + He DWD progenitor systems with intermediate orbital periods ($1.5\leq \log_{10}(P_{\rm{orb}}/\rm{day})< 2.5$) the mechanisms which impact formation efficiency are complex. The dominant way to \kb{\sout{lose} not form} CO + He DWDs is stellar mergers that occur during the second CE phase. For lower-metallicity systems, the second CE is successful, while the opposite is true for higher-metallicity systems. This is again because the envelope mass of the CE evolution donor at higher metallicity is larger than it would be at lower metallicity and thus requires more orbital energy to eject in the CE phase. This leads to shorter post-CE orbital periods after the first CE phase and hence mergers during the second phase.

@@

Long-period binaries with $\log_{10}(P_{\rm{orb}}/\rm{day})>2.5$ also display complex scenarios that hinder CO + He DWD formation. In near-equal contributions, our \cosmic\ simulations produce either stellar mergers or stable non-DWD binaries at the end of the Hubble time. At these orbital periods, stellar mergers always occur with a CE phase between a stellar companion and a CO or He WD. Similar to binaries with shorter orbital periods, the mergers occur because of increased CE donor envelope masses at higher metallicities. A subdominant channel of stable He + He DWDs can also occur when a high-metallicity primary overflows it's Roche lobe while still on the GB and thus forms a He WD. \snotes{In general, a CO + He binary fails to form because either a CO + CO or He + He is formed instead, or a merger occurs before DWD formation. We see higher formation efficiencies for FZ than F50 because CO + He DWDs prefer to form in progenitor binaries with orbital periods near $\log_{10}$(P$_{\rm{orb}}$/day)$\sim2-4$, which is where the orbital period distribution of FZ peaks.}

@@

\snotes{For our simulations with binary evolution parameter variations, we see an interplay of the physics of binary evolution with skews introduced by the metallicity-dependent orbital period distribution and binary fraction. Below, we summarize how the formation efficiency of each DWD type is impacted by the parameter variations. \kb{Two important features of our FZ and F50 models drive the differences in our binary evolution parameter variations. \sout{Of note in this analysis is that a)}}. First, the log-normal metallicity-dependent ZAMS orbital period distribution increasingly skews towards orbital periods below $10^4$ days for decreasing metallicity. However, \kb{second,} the log-uniform P$_{\rm{orb}}$ distribution of F50 produces a higher number of the shortest orbital period systems (below $\sim 100$ days) than for FZ as the tails of the FZ distribution fall quickly below this threshold.}
